The Post-2007 Declining U.S. Birth Rates imply Significant Societal Shifts in the U.S.

### 2023-02-21 outputs/paper/paper.qmd
Revised to merge
Please be mindful when commit
Revised to merge

In this paper, we gathered data from the American Economic Association, National Center for Health Statistics, and the U.S. population data. With gathered data, we generated graphs and tables to observe the trend in birth rates of different demographics, using R [@citeR]. We also uploaded and cleaned data using R [@citeR], tidyr [@citetidyr], dplyr [@citedplyr], tidyverse [@citetidyverse], and haven [@citehaven] packages. Figures and tables were created with knitr [@citeknit], ggplot2 [@citeggplot], kableExtra [@citekable], and dplyr [@citedplyr].

@fig-1 depicts U.S. birth rate trends over the last 40 years. In Figure 1, we can see that the trend is fluctuating relatively stable. However, during the post-2007 recession, it takes a steep decline with no sign of turning back.

In @sec-first-point, we discuss how decomposing the decline in the birth rate between demographic groups helps us to have a clearer picture of the topic. As can be seen in @fig-2 and @fig-3

@@

We conclude the paper by raising a concern that this societal phenomenon imposes some severe societal consequences and, therefore, should be considered at a high level. Most importantly, we need to consider what it means for the nation's well-being when there are no signs of returning from the negative trend in birth rates. 
\newpage

@@

## Weaknesses and next steps

\appendix

# Appendix {.unnumbered}

# Additional details
## A Potential Explanation: Shifting Priorities {#sec-third-point}

>>>>

In this paper, we gathered data from the American Economic Association, National Center for Health Statistics, and the U.S. population data. With the gathered data, we generated graphs and tables to observe the trend in birth rates of different demographics using R [@citeR]. We also uploaded and cleaned data using R [@citeR], tidyr [@citetidyr], dplyr [@citedplyr], tidyverse [@citetidyverse], and haven [@citehaven] packages. Figures and tables were created with knitr [@citeknit], ggplot2 [@citeggplot], kableExtra [@citekable], and dplyr [@citedplyr].

@fig-1 depicts U.S. birth rate trends over the last 40 years. Here, we can see that the trend is fluctuating relatively stable. However, during the post-2007 recession, it takes a steep decline with no sign of returning. We start our journey to explore potential explanations for the decline.

In @sec-first-point, we discuss how decomposing the decline in the birth rate between demographic groups offers us a clearer picture of the topic. As shown in @fig-2 and @fig-3, by dissecting the demographics into age groups and ethnicity groups, we notice that there is a significant amount of decrease among the white ethnicity women between ages 15 and 24, while other ethnic groups are showing a minor decrease.

@@

We conclude the paper by raising a concern that this societal phenomenon imposes some severe societal consequences and, therefore, should be considered at a high level. Most importantly, we need to consider what it means for the nation's well-being when there are no signs of returning from the negative trend in birth rates.

@@
